Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint

### 2020-01-06 analysis/VOCC/2-biomass-depth-only-model.Rmd
update checks on biomass models

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Redbanded Rockfish"
    #species: "Pacific Ocean Perch"
    #species: "Pacific Cod"
    #species: "Petrale Sole"
    species: "Pacific Halibut"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
---

@@

# Build spatiotemporal density model
Load and filter data 
```{r}
biomass <- readRDS(paste0(
  "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
))
covars <- readRDS("data/event-covariates.rds")
data <- dplyr::left_join(biomass, covars)

@@

    adult_biomass <- sdmTMB::sdmTMB(
      data = data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      # ar1_fields = TRUE,
      include_spatial = TRUE,
      # reml = TRUE,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

@@

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      # ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
      reml = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0(
        "data/", spp,
        "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
      )
    )
  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

Save predictions for full spatial grid

```{r}
# fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (any(names(data) == "adult_density")) {
  # if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))

@@

  predictions <- predict(total_biomass, newdata = nd)
  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
  ))
}
```

```{r}
try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

@@

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
))
```

@@

### Build time-varying depth plots
#### Model without environmental covariates
```{r }
max_depth_found <- max(data[data$present == 1, ]$raw_depth, na.rm = TRUE)
rm(depth_model_list)
rm(adult_biomass)
rm(imm_biomass)
rm(total_biomass)

@@

d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  if (depth_model_list[[i]]$model$convergence == 0) {
    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")

    if (length(d[[i]]) == 0) {
      depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
    } else {
      d[[i]]$x <- exp(d[[i]]$x)
      depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
        ggtitle(paste(species, names(depth_model_list[i])))
    }
  } else {
    depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  }
}
print(depth_plots)
```

@@

>>>>

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with time-varying depth and trawl footprint"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Redbanded Rockfish"
    #species: "Pacific Ocean Perch"
    #species: "Pacific Cod"
    #species: "Petrale Sole"
    #species: "Widow Rockfish"
    species: "Bocaccio"
    region: "All synoptic surveys"
    knots: 500
    covariates: "" # +as.factor(ssid)
    #additional fixed-effect covariates
    #covs: "-ssid-only" # covariate string used to name model rds 
    covs: "-tv-depth-only"
    #covs: "-trawled-ssid"
    update_model: FALSE
    update_predictions: FALSE
    update_model_check: FALSE
---

@@

# Build spatiotemporal density model
Load and filter data 
```{r}
biomass <- readRDS(paste0(
  "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
))

if(nrow(biomass)<4000) {stop("Need to recalculate split by maturity!")}

covars <- readRDS("data/event-covariates.rds")
data <- dplyr::left_join(biomass, covars)

@@

    adult_biomass <- sdmTMB::sdmTMB(
      data = data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE,
      include_spatial = TRUE,
      # reml = TRUE,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

@@

   try({ imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
    ))

    imm_biomass <- sdmTMB::sdmTMB(
      data = data,
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, #+ trawled
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE, # changed to TRUE on Oct 17 2019
      # reml = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass,
      file = paste0("data/", spp, "/mod-imm-biomass-", 
        spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
   })

  } else {
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

@@

Save predictions for full spatial grid
```{r}
# fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)
if(params$update_predictions) {
if (any(names(data) == "adult_density")) {
  # if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))

@@

  predictions <- predict(total_biomass, newdata = nd)
  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
  ))
}
}
```

```{r}

if(params$update_model_check) {
try({
  biomass <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

@@

saveRDS(point_predictions, file = paste0(
  "data/", spp,
  "/check-mod-predictions-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
))
}
```

@@

### Build time-varying depth plots
#### Model without environmental covariates
```{r }
max_depth_found <- max(data[data$present == 1, ]$raw_depth, na.rm = TRUE)
rm(depth_model_list)
rm(adult_biomass)
rm(imm_biomass)
rm(total_biomass)
rm(depth_plots)
#max_depth_found <- 800

@@

d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  #if (depth_model_list[[i]]$model$convergence == 0) {
    d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")

    if (length(d[[i]]) == 0) {
      depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
    } else {
      d[[i]]$x <- exp(d[[i]]$x)
      depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth (without environmental variables)", xlimits = c(0, max_depth_found)) +
        ggtitle(paste(species, names(depth_model_list[i])))
    }
  # } else {
  #   depth_plots[[i]] <- grid::grid.rect(gp = grid::gpar(col = "white"))
  # }
}
print(depth_plots)
```

@@

```{r}
sdmTMB:::get_convergence_diagnostics(depth_model_list[[1]])
```
